---
title: "Practical Machine Learning - Course Project: Writeup"
author: "Intan Ismail"
date: "Saturday, September 19, 2015"
output: html_document
---

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible 
to collect a large amount of data about personal activity relatively 
inexpensively. These type of devices are part of the quantified self movement –
a group of enthusiasts who take measurements about themselves regularly to 
improve their health, to find patterns in their behavior, or because they are 
tech geeks. One thing that people regularly do is quantify how much of a 
particular activity they do, but they rarely quantify how well they do it. 

In this project, we will be to use data from accelerometers on the belt, 
forearm, arm, and dumbell of 6 participants. They were asked to perform barbell 
lifts correctly and incorrectly in 5 different ways. More information is 
available from the website here: http://groupware.les.inf.puc-rio.br/har 
(see the section on the Weight Lifting Exercise Dataset). 

### Data 

The training data for this project are downloaded from here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are downloaded from here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: 
http://groupware.les.inf.puc-rio.br/har. 

### Loading and Processing Data 

#### Load Weight Lifting Exercises Dataset

```{r,cache=TRUE, warning=FALSE}
# load the data
train_ds <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
test_ds <- read.csv("pml-testing.csv", na.strings = c("NA", "#DIV/0!", ""))

# cleanup the data
train_ds <- train_ds[, colSums(is.na(train_ds)) == 0]
test_ds <- test_ds[, colSums(is.na(test_ds)) == 0]
```

```{r}
# get train dataset dimension
dim(train_ds)
```

The train dataset has 19622 observations and 60 variables.

```{r}
# get test dataset dimension
dim(test_ds)
```

The test dataset has 20 observations and 60 variables.

#### Partition train dataset for analysis

```{r}
library(caret)
set.seed(12345)
inTrain <- createDataPartition(y = train_ds$classe,
                             p = 0.7,
                             list = FALSE)
training <- train_ds[inTrain,]
testing <- train_ds[-inTrain,]
```

#### Model Selection

We use randon forest as our model for selection. From Wikipedia, Random forests 
are an ensemble learning method for classification, regression and other tasks, 
that operate by constructing a multitude of decision trees at training time and 
outputting the class that is the mode of the classes (classification) or mean 
prediction (regression) of the individual trees. Based on this and from accuracy 
stand point this model will fit as a model for this project. 

```{r}
model_rf <- train(classe ~ ., data = training, method = "rf",
                  tuneGrid = data.frame(mtry = 3),
                  trControl = trainControl(method = "none"))
model_rf
```

#### Model Validation

We now test our model on the training and testing sets and perform cross 
validatation.

```{r}
# Training set
training_set <- predict(model_rf, training)
print(confusionMatrix(training_set, training$classe))
```

Our model performs with highest accuracy against the training set but we need
to cross validate against testing set to see if we have avoided overfitting. 

```{r}
# Testing set
testing_set <- predict(model_rf, testing)
print(confusionMatrix(testing_set, testing$classe))
```

The testing set accuracy is 99% with sample error of 1% which shows our model
performs rather good.

#### Test set prediction

We now test the prediction of our algorithm on the test set.

```{r}
test_set <- predict(model_rf, test_ds)
test_set
```

We then save and write the output to post it to the Course Project: Submission. 

```{r}
answers <- as.vector(test_set)

pml_write_files = function(x) {
    n = length(x)
    for(i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], 
                    file = filename, 
                    quote = FALSE, 
                    row.names = FALSE, 
                    col.names = FALSE)
    }
}

pml_write_files(answers)
```